Primal-dual algorithm for computing Nash equilibria in sequential games with imcomplete information

### 2014-12-17 nips2015.tex
wip

$a(t)$ is the payoff received by player $1$ at node $t$. That the game is zero-sum means that player 2 receives a payoff of $-a(t)$. Note that $a(t)$ is $0$ for all non-leaf nodes.

@@

Each point $z \in Q_k$ are called \textit{realization plan} and induces a \textit{behavioural-strategy} $\beta_k$ such that for each information set $h \in \mathcal{I}_k$ the probability for player $k$ to make a choice $c \in \mathcal{C}_h$ from any node in $h$ is
\begin{equation}
  \beta_k(h, c) := \begin{cases}
    z(\sigma_hc)/z(\sigma_h), é\mbox{if } z(\sigma_h) > 0\\
    0, &\mbox{Otherwise}.
  \end{cases}
\end{equation}

@@

@@

\paragraph{Example of Nash equilibirum}
As an illustration, the pair $(x^*, y^*)$ given by
$x^* = [1, .478, .522, .174, .826]^T$ and
$y^* = [1, 1/2, 1/2]^T$ is a Nash equilibrium for the sequence-form game given by (not showing zero entries)\\
$A = \left[\begin{array}{ccc}
  &   &  \\
  &   &  \\
  & 1 & -1\\
  & -2 & 4\\
1 &   &  
\end{array}\right]$, $E_1 = \left[\begin{array}{ccccc}
  1 &   &   &   &  \\
  -1 & 1 & 1 &   &  \\
  -1 &   &   & 1 & 1
\end{array}\right]$,
$E_2 = \left[\begin{array}{ccc}
  1 &   &  \\
  -1 & 1 & 1
\end{array}\right]$, $e_1 = [1, 0, 0]^T$, and $e_2 = [1, 0]^T$.

@@

Note that $i_{C \cap D} = i_C + i_D$. The \textit{euclidean projection operator} onto $C$, denoted $\Pi_C$ is the function
$\Pi: \mathbb{R}^n \mapsto C$, which maps a point $x \in \mathbb{R}^n$ to the (necessarily unique) point $\Pi_C(x)$ of $C$ which is closed to $x$. Precisely,
\begin{equation}
  \Pi_C(x) := \underset{c \in C}{argimin}\text{ }\|c - x\|^2
\end{equation}
For example, $\Pi_{\mathbb{R}^n_+}(x) = (x)_+, \forall x \in \mathbb{R}^n$.

@@

%% Theorem \ref{thm:pd}, can be interpreted as follows: 2 players $x$ (player 1)  and $y$ (player 2) go heads-on in a zero-sum game with payoff matrix $A$. Their strategy, profiles are $Q_1$ and $Q_2$ respectively, meaning that if player $j$ players a strategy which is outside $Q_k$ then they are panalized by inflicting an infinite loss yupon them. To remedy this, they recruit assistants $u$ and $v$, whose sole goal is to preven the ...
\begin{proof} First observe that $\forall (x, y) \in \mathbb{R}^{n_1} \times \mathbb{R}^{n_2}$, we have
  \begin{eqnarray*}
   -i_{Q_1}(x) = -i_{\mathbb{R}^{n_1}_+}(x) + \underset{v \in \mathbb{R}^{p_1}}{min}\text{}{v^T(e_1 - E_1x)},
      \end{eqnarray*}
  and
\begin{eqnarray*}
  i_{Q_2}(y) = i_{\mathbb{R}^{n_2}_+}(y) + \underset{u \in \mathbb{R}^{p_2}}{max}\text{}{u^T(E_2y - e_2)}
\end{eqnarray*}
  Thus,% problem \eqref{eq:opt_pb} is equivalent to
  \begin{eqnarray*}
    \begin{split}
  \underset{y \in Q_2}{min}\text{ }\underset{x \in Q_1}{max}\text{ }{x^TAy} &=
  \underset{y \in \mathbb{R}^{n_2}}{min}\text{ }\left(\underset{x \in \mathbb{R}^{n_1}}{max}\text{ }
x^TAy - i_{Q_1}(x)\right) + i_{Q_2}(y) \\
&= \underset{y \in \mathbb{R}^{n_2}}{min}\text{ }\underset{x \in \mathbb{R}^{n_1}}{max}\text{ }x^TAy -i_{\mathbb{R}^{n_1}_+}(x) + \underset{v \in \mathbb{R}^{p_1}}{min}\text{}{v^T(e_1 - E_1x)} + i_{\mathbb{R}^{n_2}_+}(y) + \underset{u \in \mathbb{R}^{p_2}}{max}\text{}{u^T(E_2y - e_2)}\\
      &= \underset{y \in \mathbb{R}^{n_2}}{min}\text{ }\underset{x \in \mathbb{R}^{n_1}}{max}\text{ }\underset{v \in \mathbb{R}^{p_1}}{min}\text{ }\underset{u \in \mathbb{R}^{p_2}}{min}\text{ }x^TAy - x^TE_1^Tv + u^TE_2y + i_{\mathbb{R}^{n_2}_+}(y) + e_1^Tv -(i_{\mathbb{R}^{n_1}_+}(x) + e_2^Tu)\\
      &= \underset{y \in \mathbb{R}^{n_2}}{min}\text{ }\underset{x \in \mathbb{R}^{n_1}}{max}\text{ } \underset{v\in \mathbb{R}^{p_1}}{min}\text{ }\underset{u \in \mathbb{R}^{p_2}}{max}
      {\begin{bmatrix}x\\u\end{bmatrix}^TK\begin{bmatrix}y\\v\end{bmatrix} + G(y, v) - H(x, u)}\\
      &= \underset{y \in \mathbb{R}^{n_2}}{min}\text{ }\underset{v\in \mathbb{R}^{p_1}}{min}\text{ }\underset{x \in \mathbb{R}^{n_1}}{max}\text{ }
\underset{u \in \mathbb{R}^{p_2}}{max}
      {\begin{bmatrix}x\\u\end{bmatrix}^TK\begin{bmatrix}y\\v\end{bmatrix} + G(y, v) - H(x, u)}\\
      &=     \underset{y \in \mathbb{R}^{n_2}, v\in \mathbb{R}^{p_1}}{min}\text{ }\underset{x \in \mathbb{R}^{n_1}, u \in \mathbb{R}^{p_2}}{max}
           {\begin{bmatrix}x\\u\end{bmatrix}^TK\begin{bmatrix}y\\v\end{bmatrix} + G(y, v) - H(x, u)}
    \end{split}
  \end{eqnarray*}
where $K$, $G$, and $H$ are as given in \eqref{eq:things}. Thus problem \eqref{eq:opt_pb} is equivalent to problem \eqref{eq:unconstrained_pb} as claimed.

@@

\section{Application to Poker: Kuhn Poker}
\label{sec:kuhn}
The Kuhn 3-card Poker has sequence-form specification given by (not showing zero entries)\\
$A = \left[\begin{array}{ccccccccccccc}
  &   &   &   &   &   &   &   &   &   &   &   &  \\
  &   &   &   &   &   &   & -1 / 6 &   &   &   & -1 / 6 &  \\
  &   &   &   &   &   &   &   & -1 / 6 &   &   &   & -1 / 6\\
  &   &   &   &   &   &   &   & -1 / 3 &   &   &   & -1 / 3\\
  &   &   &   &   & 1 / 6 & -1 / 3 &   &   & 1 / 6 & -1 / 3 &   &  \\
  &   &   & 1 / 6 &   &   &   &   &   &   &   & -1 / 6 &  \\
  &   &   &   & -1 / 6 &   &   &   &   &   &   &   & -1 / 6\\
  &   &   &   & 1 / 3 &   &   &   &   &   &   &   & -1 / 3\\
  & 1 / 6 & 1 / 3 &   &   &   &   &   &   & 1 / 6 & -1 / 3 &   &  \\
  &   &   & 1 / 6 &   &   &   & 1 / 6 &   &   &   &   &  \\
  &   &   &   & -1 / 6 &   &   &   & -1 / 6 &   &   &   &  \\
  &   &   &   & 1 / 3 &   &   &   & 1 / 3 &   &   &   &  \\
  & 1 / 6 & 1 / 3 &   &   & 1 / 6 & 1 / 3 &   &   &   &   &   &  
\end{array}\right]$,\\
$E_1 = \left[\begin{array}{ccccccccccccc}
1 &   &   &   &   &   &   &   &   &   &   &   &  \\
-1 &   &   &   &   &   &   &   &   & 1 &   &   & 1\\
-1 & 1 &   &   & 1 &   &   &   &   &   &   &   &  \\
-1 &   &   &   &   & 1 &   &   & 1 &   &   &   &  \\
  & -1 & 1 & 1 &   &   &   &   &   &   &   &   &  \\
  &   &   &   &   & -1 & 1 & 1 &   &   &   &   &  \\
  &   &   &   &   &   &   &   &   & -1 & 1 & 1 &  
\end{array}\right]$, $e_1 = e_2 = [1, 0, 0, 0, 0, 0, 0]^T$,\\
and $E_2 = \left[\begin{array}{ccccccccccccc}
1 &   &   &   &   &   &   &   &   &   &   &   &  \\
-1 &   &   &   &   &   &   & 1 & 1 &   &   &   &  \\
-1 &   &   &   &   &   &   &   &   & 1 & 1 &   &  \\
-1 &   &   &   &   & 1 & 1 &   &   &   &   &   &  \\
-1 &   &   &   &   &   &   &   &   &   &   & 1 & 1\\
-1 & 1 & 1 &   &   &   &   &   &   &   &   &   &  \\
-1 &   &   & 1 & 1 &   &   &   &   &   &   &   &  
\end{array}\right]$.\\
The pair $(x^*, y^*)$ of realization plans given by\\
$x^* = [1, 0.759, 0.759, 0, 0.241, 1, 0.425, 0.575, 0, 0.275, 0, 0.275, 0.725]^T$ and
$y^* = [1, 1, 0, 0.667, 0.333, 0.667, 0.333, 1, 0, 0, 1, 0, 1]^T$ is a Nash equlibrium for the game, computed using Algorithm  \ref{Tab:algo_simplified}. The convergence curve is shown in Fig \ref{Tab:conv_curves}. One easy checks that this equilibrium is feasible. Indeed,  $E_1x^* - e_1 = [4.76 \times 10^{-5}, -1.91 \times 10^{-5}, 5.67 \times 10^{-5}, 8.23 \times 10^{-6}, 2.90 \times 10^{-5}, -8.62 \times 10^{-7}, -1.96 \times 10^{-5}]^T$ and $E_2y^* - e_2 = [-7.04 \times 10^{-7}, 2.27 \times 10^{-6}, -3.29 \times 10^{-6}, -1.50 \times 10^{-6}, 2.92 \times 10^{-6}, -4.97 \times 10^{-7}, -5.85 \times 10^{-7}]^T$. Finally, $x^*TAy^* = -0.055593685705289997$, which agrees to 4 d.p with the value of $-1 / 18$ computed analytically by Kuhn in his 1954 paper.

>>>>

$a(t)$ is the \textit{payoff} received by player $1$ at node $t$. That the game is zero-sum means that player 2
receives a payoff of $-a(t)$. Note that $a(t)$ is $0$ for all non-leaf nodes, since the game must continue.
The goal of a \textit{rational} player is that the game should end at a node for which her payoff is as large as possible.
We will assume implicitly that each (non-chance) player is rational.

@@

Each point $z \in Q_k$ is called a \textit{realization plan} and induces a \textit{behavioural-strategy} $\beta_k$ such that for each information set $h \in \mathcal{I}_k$ and each choice $c \in C_h$, the probability for player $k$ to make a choice $c \in \mathcal{C}_h$ from any node in $h$ is
\begin{equation}
  \beta_k(h, c) := \begin{cases}
    z(\sigma_hc)/z(\sigma_h), é\mbox{if } z(\sigma_h) > 0\\
    0, &\mbox{Otherwise}.
  \end{cases}
\end{equation}

@@

\paragraph{Example: Sequence-form representation of Kuhn Poker}
\label{sec:kuhn_sf}
The Kuhn 3-card Poker has sequence-form specification given by (not showing zero entries)\\
$A = \left[\begin{array}{ccccccccccccc}
  &   &   &   &   &   &   &   &   &   &   &   &  \\
  &   &   &   &   &   &   & -1 / 6 &   &   &   & -1 / 6 &  \\
  &   &   &   &   &   &   &   & -1 / 6 &   &   &   & -1 / 6\\
  &   &   &   &   &   &   &   & -1 / 3 &   &   &   & -1 / 3\\
  &   &   &   &   & 1 / 6 & -1 / 3 &   &   & 1 / 6 & -1 / 3 &   &  \\
  &   &   & 1 / 6 &   &   &   &   &   &   &   & -1 / 6 &  \\
  &   &   &   & -1 / 6 &   &   &   &   &   &   &   & -1 / 6\\
  &   &   &   & 1 / 3 &   &   &   &   &   &   &   & -1 / 3\\
  & 1 / 6 & 1 / 3 &   &   &   &   &   &   & 1 / 6 & -1 / 3 &   &  \\
  &   &   & 1 / 6 &   &   &   & 1 / 6 &   &   &   &   &  \\
  &   &   &   & -1 / 6 &   &   &   & -1 / 6 &   &   &   &  \\
  &   &   &   & 1 / 3 &   &   &   & 1 / 3 &   &   &   &  \\
  & 1 / 6 & 1 / 3 &   &   & 1 / 6 & 1 / 3 &   &   &   &   &   &  
\end{array}\right]$,\\
$E_1 = \left[\begin{array}{ccccccccccccc}
1 &   &   &   &   &   &   &   &   &   &   &   &  \\
-1 &   &   &   &   &   &   &   &   & 1 &   &   & 1\\
-1 & 1 &   &   & 1 &   &   &   &   &   &   &   &  \\
-1 &   &   &   &   & 1 &   &   & 1 &   &   &   &  \\
  & -1 & 1 & 1 &   &   &   &   &   &   &   &   &  \\
  &   &   &   &   & -1 & 1 & 1 &   &   &   &   &  \\
  &   &   &   &   &   &   &   &   & -1 & 1 & 1 &  
\end{array}\right]$, $e_1 = e_2 = [1, 0, 0, 0, 0, 0, 0]^T$,\\
and $E_2 = \left[\begin{array}{ccccccccccccc}
1 &   &   &   &   &   &   &   &   &   &   &   &  \\
-1 &   &   &   &   &   &   & 1 & 1 &   &   &   &  \\
-1 &   &   &   &   &   &   &   &   & 1 & 1 &   &  \\
-1 &   &   &   &   & 1 & 1 &   &   &   &   &   &  \\
-1 &   &   &   &   &   &   &   &   &   &   & 1 & 1\\
-1 & 1 & 1 &   &   &   &   &   &   &   &   &   &  \\
-1 &   &   & 1 & 1 &   &   &   &   &   &   &   &  
\end{array}\right]$.\\

@@

%% \paragraph{Example of Nash equilibirum}
%% As an illustration, the pair $(x^*, y^*)$ given by
%% $x^* = [1, .478, .522, .174, .826]^T$ and
%% $y^* = [1, 1/2, 1/2]^T$ is a Nash equilibrium for the sequence-form game given by (not showing zero entries)\\
%% $A = \left[\begin{array}{ccc}
%%   &   &  \\
%%   &   &  \\
%%   & 1 & -1\\
%%   & -2 & 4\\
%% 1 &   &  
%% \end{array}\right]$, $E_1 = \left[\begin{array}{ccccc}
%%   1 &   &   &   &  \\
%%   -1 & 1 & 1 &   &  \\
%%   -1 &   &   & 1 & 1
%% \end{array}\right]$,
%% $E_2 = \left[\begin{array}{ccc}
%%   1 &   &  \\
%%   -1 & 1 & 1
%% \end{array}\right]$, $e_1 = [1, 0, 0]^T$, and $e_2 = [1, 0]^T$.

@@

Note that $i_{C \cap D} = i_C + i_D$. The \textit{euclidean projection operator} onto $C$, denoted $\Pi_C$ is the function
$\Pi: \mathbb{R}^n \mapsto C$, which maps a point $x \in \mathbb{R}^n$ to the (necessarily unique) point $\Pi_C(x)$ of $C$ which is closed to $x$. Precisely,
\begin{equation}
  \Pi_C(x) := \underset{c \in C}{argimin}\text{ }\|c - x\|^2
\end{equation}
For example, the metric projections onto the nonnegative orthant and the closed unit-disk $\bar{\Delta}$
% := \{x|\|x\| \le 1\}$
are $\Pi_{\mathbb{R}^n_+}(x) \equiv (x)_+$ and $\Pi_{\bar{\Delta}}(z) \equiv \dfrac{z}{max(\|z\|, 1)}$ respectively.

@@

%% Theorem \ref{thm:pd}, can be interpreted as follows: 2 players $x$ (player 1)  and $y$ (player 2) go heads-on in a zero-sum game with payoff matrix $A$. Their strategy, profiles are $Q_1$ and $Q_2$ respectively, meaning that if player $j$ players a strategy which is outside $Q_k$ then they are panalized by inflicting an infinite loss yupon them. To remedy this, they recruit assistants $u$ and $v$, whose sole goal is to preven the ...
\begin{proof} First observe that $\forall (x, y) \in \mathbb{R}^{n_1} \times \mathbb{R}^{n_2}$, we have
  \begin{eqnarray}
    \left.
    \begin{split}
      -i_{Q_1}(x) = -i_{\mathbb{R}^{n_1}_+}(x) + \underset{v \in \mathbb{R}^{p_1}}{min}\text{}{v^T(e_1 - E_1x)}\\
      i_{Q_2}(y) = i_{\mathbb{R}^{n_2}_+}(y) + \underset{u \in \mathbb{R}^{p_2}}{max}\text{}{u^T(E_2y - e_2)}
      \end{split}
    \right\}
  \end{eqnarray}
  Thus,% problem \eqref{eq:opt_pb} is equivalent to
  \begin{eqnarray*}
    \begin{split}
  \underset{y \in Q_2}{min}\text{ }\underset{x \in Q_1}{max}\text{ }{x^TAy} &=
  \underset{y \in \mathbb{R}^{n_2}}{min}\text{ }\left(\underset{x \in \mathbb{R}^{n_1}}{max}\text{ }
x^TAy - i_{Q_1}(x)\right) + i_{Q_2}(y) \\
&= \underset{y \in \mathbb{R}^{n_2}}{min}\text{ }\underset{x \in \mathbb{R}^{n_1}}{max}\text{ }x^TAy -i_{\mathbb{R}^{n_1}_+}(x) + \underset{v \in \mathbb{R}^{p_1}}{min}\text{}{v^T(e_1 - E_1x)} + i_{\mathbb{R}^{n_2}_+}(y) + \underset{u \in \mathbb{R}^{p_2}}{max}\text{}{u^T(E_2y - e_2)}\\
      &= \underset{y \in \mathbb{R}^{n_2}}{min}\text{ }\underset{x \in \mathbb{R}^{n_1}}{max}\text{ }\underset{v \in \mathbb{R}^{p_1}}{min}\text{ }\underset{u \in \mathbb{R}^{p_2}}{min}\text{ }x^TAy - x^TE_1^Tv + u^TE_2y + i_{\mathbb{R}^{n_2}_+}(y) + e_1^Tv -(i_{\mathbb{R}^{n_1}_+}(x) + e_2^Tu)\\
      &= \underset{y \in \mathbb{R}^{n_2}}{min}\text{ }\underset{x \in \mathbb{R}^{n_1}}{max}\text{ } \underset{v\in \mathbb{R}^{p_1}}{min}\text{ }\underset{u \in \mathbb{R}^{p_2}}{max}
      {\begin{bmatrix}x\\u\end{bmatrix}^TK\begin{bmatrix}y\\v\end{bmatrix} + G(y, v) - H(x, u)}\\
      &= \underset{y \in \mathbb{R}^{n_2}}{min}\text{ }\underset{v\in \mathbb{R}^{p_1}}{min}\text{ }\underset{x \in \mathbb{R}^{n_1}}{max}\text{ }
\underset{u \in \mathbb{R}^{p_2}}{max}
      {\begin{bmatrix}x\\u\end{bmatrix}^TK\begin{bmatrix}y\\v\end{bmatrix} + G(y, v) - H(x, u)}\\
      &=     \underset{y \in \mathbb{R}^{n_2}, v\in \mathbb{R}^{p_1}}{min}\text{ }\underset{x \in \mathbb{R}^{n_1}, u \in \mathbb{R}^{p_2}}{max}
           {\begin{bmatrix}x\\u\end{bmatrix}^TK\begin{bmatrix}y\\v\end{bmatrix} + G(y, v) - H(x, u)}
    \end{split}
  \end{eqnarray*}
where $K$, $G$, and $H$ are as given in \eqref{eq:things}. Thus problem \eqref{eq:opt_pb} is equivalent to problem \eqref{eq:unconstrained_pb} as claimed.

@@

\section{Application to Poker: Kuhn Poker}
\label{sec:kuhn}
The pair $(x^*, y^*)$ of realization plans given by\\
$x^* = [1, 0.759, 0.759, 0, 0.241, 1, 0.425, 0.575, 0, 0.275, 0, 0.275, 0.725]^T$ and
$y^* = [1, 1, 0, 0.667, 0.333, 0.667, 0.333, 1, 0, 0, 1, 0, 1]^T$ is a Nash equlibrium for the Kuhn Poker presented in \ref{sec:kuhn_sf}, computed using Algorithm  \ref{Tab:algo_simplified}. The convergence curve is shown in Fig \ref{Tab:conv_curves}. One easy checks that this equilibrium is feasible. Indeed,  $E_1x^* - e_1 = [4.76 \times 10^{-5}, -1.91 \times 10^{-5}, 5.67 \times 10^{-5}, 8.23 \times 10^{-6}, 2.90 \times 10^{-5}, -8.62 \times 10^{-7}, -1.96 \times 10^{-5}]^T$ and $E_2y^* - e_2 = [-7.04 \times 10^{-7}, 2.27 \times 10^{-6}, -3.29 \times 10^{-6}, -1.50 \times 10^{-6}, 2.92 \times 10^{-6}, -4.97 \times 10^{-7}, -5.85 \times 10^{-7}]^T$. Finally, $x^*TAy^* = -0.055593685705289997$, which agrees to 4 d.p with the value of $-1 / 18$ computed analytically by Kuhn in his 1954 paper.
